\myTitle

### 2015-02-06 tex/main-body-text.tex
couple new figures. troubleshooting reproducability on another computer

\section*{Comparison of {\sc structure} and {\sc flock}}
The implementation of {\sc flock} as a VBA procedure (macro) in Microsoft Excel precludes
a rigorous evaluation of its behavior due to the inability to efficiently batch process. Consequenty, 
we adapted the source code from {\sc structure} as describe above to create {\sc flockture}
(code provided on GitHub repository https://github.com/eriqande/flockture) in order to accurately
assess the strengths and weaknesses of the {\sc flock} algorithm. First, using simulated 
data we demonstrated that {\sc flock} and {\sc flockture} are essentially equivalent programs.
Second, we simulated and analyzed datasets with {\sc flockture} and three models in {\sc structure} 
(the admixture and correlated allele frequency (A-C) model, 
the no-admixture and correlated allele frequency (NA-C) model, 
and the no-admixture and non-correlated allele frequency (NA-NC) model) to help better understand
under what conditions, if any, running {\sc flock} may be preferable to {\sc structure} .

For each set of runs, {\sc flock} reports the 
normalized likelihood values (the exponentiated log-likelihood value for an individual's 
membership to each of $K$ clusters, scaled to sum to one) for a run belonging to the 
longest plateau - refered to as the 'best run' by \citep{Duc&Tur2012} (see below).
These values can be compared to the $q_i$ values
returned by {\sc structure}, which, in the no-admixture case, are posterior 
probabilities of cluster membership.\citet{Duc&Tur2012} describe a method 
for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must specify both the number of runs (different initial allocations of individuals 
to each cluster) as well as the number of iterations. (In each iteration every individual 
has the opportunity to be reallocated 
to a different cluster.) At the conclusion of each run, the log-likelihood value for 
membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the cluster with the highest likelihood 
and second highest likelihood, is calculated for each individual. {\sc flock} identifies runs that 
have converged to the same solution by detecting that they have the same mean LLOD score, and, in 
{\sc flock} parlance, a series of runs that have converged to the same solution are termed a ``plateau.'' 
The plateau record is a list of how many runs converged to the same allocation
of individuals. \citet{Duc&Tur2012} developed two stopping rules that they use to estimate $K$. 
\citet{Duc&Tur2012} advocate running  {\sc flock} until either a single plateau of
$\geq 6$ runs is reached or four consecutive $K$ values yield no plateaus. If the first stopping condition is satisfied,
the $K$ for which there is a single plateau of $\geq 6$ runs is the point estimate for $K$. Alternatively, 
if multiple plateaus are observed with at least one plateau of $\geq 6$ runs, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. If no plateaus of $\geq 6$ runs are observed, $K$
is declared {\em undecided}.

@@

Once we were confident that no major differences in inference existed between  {\sc flock} and {\sc flockture}
we used  {\sc flockture} to analyze a many more datasets that encompassed a broader 
range of population differentiation. Fourteen datasets were simulated as above (five populations 
comprised of 250, 250, 400, 550, and 550 individuals with 15 loci genotypes)
with average pairwise $F_{ST}$ values ranging between 0.11 and 0.005. 
 {\sc flockture} was run with the default run parameters and each run of 
{\sc structure} consisted of a 5,000 sweeep burn-in followed by a 20,000 sweep sample
from the posterior distribution. All runs were performed on a Macbook with a 2.3 GHz i7 processor
and 16GB of RAM running OS X ver. 10.7.5. The program {\sc clumpp} \citep{Jak&Ros2007} 
was used to relabel clusters among different runs,
while the program {\sc district} \citep{Rosenberg2004} was used to plot individual \textit{$q_i$} values.

@@

 \begin{figure*}
\centering
  \includegraphics[width=.9\linewidth]{images/Figures-Pat/FvSloss1.pdf}
  \caption{Zero-one loss over 9 replicate runs of the admixture and correlated allele 
  frequency (A-C) model, 
the no-admixture and correlated allele frequency (NA-C) model, 
and the no-admixture and non-correlated allele frequency (NA-NC) model in {\sc structure} and {\sc flockture}. 
Average pairwise $F_{ST}$ 
values between populations is listed above each graph.}
  \label{fig:FvSloss}
\end{figure*} % These graphs need to be redone so that text is legible and corrected for the NA-C NOT the A-NC 
%model in structure. I probably had a transcription error in R. Check code to be sure.

>>>>

\section*{Comparison of {\sc structure} and {\sc flock}}
{\sc flock}, initially designed to map admixture in hybrid populations 
\citep{Duc&Tur2009}, was recently proposed as an alternative to 
{\sc structure} to estimate the number of populations from a collection with 
unsampled source populations \citep{Duc&Tur20123}. In doing so, 
{\sc flock} relies on individuals with like genotypes clustering together 
through an iterative process. It stands to reason then, a reasonable way 
to gauge how well the algorithm accomplishes the main objective 
clustering of individuals into correct groups is to simulate data of known 
population structure with varying levels of differentiation and 
evaluate how well it does assuming the correct number of populations. 
The implementation of {\sc flock} as a VBA procedure (macro) in 
Microsoft Excel precludes a rigorous evaluation of its behavior due 
to the inability to efficiently batch process. Consequenty, we adapted 
the source code from {\sc structure} as describe above to create {\sc flockture}
(code provided on GitHub repository https://github.com/eriqande/flockture) 
in order to accurately assess the strengths and weaknesses of the {\sc flock} 
algorithm. First, using simulated data we demonstrated that {\sc flock} 
and {\sc flockture} are essentially equivalent programs.
Second, we simulated and analyzed datasets with {\sc flockture} 
and three models in {\sc structure} (the admixture and correlated 
allele frequency (A-C) model, the no-admixture and correlated 
allele frequency (NA-C) model, and the no-admixture and non-correlated 
allele frequency (NA-NC) model) to help better understand
under what conditions, if any, running {\sc flock} may be preferable to {\sc structure} .

For each set of runs, {\sc flock} reports the 
normalized likelihood values (the exponentiated log-likelihood value for an individual's 
membership to each of $K$ clusters, scaled to sum to one) for a run belonging to the 
longest plateau - refered to as the 'best run' by \citep{Duc&Tur2012} (see below).
These values can be compared to the $q_i$ values
returned by {\sc structure}, which, in the no-admixture case, are posterior 
probabilities of cluster membership. \citet{Duc&Tur2012} describe a method 
for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must specify both the number of runs (different initial allocations of individuals 
to each cluster) as well as the number of iterations. (In each iteration every individual 
has the opportunity to be reallocated 
to a different cluster.) At the conclusion of each run, the log-likelihood value for 
membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the cluster with the highest likelihood 
and second highest likelihood, is calculated for each individual. {\sc flock} identifies runs that 
have converged to the same solution by detecting that they have the same mean LLOD score, and, in 
{\sc flock} parlance, a series of runs that have converged to the same solution are termed a ``plateau.'' 
The plateau record is a list of how many runs converged to the same allocation
of individuals. \citet{Duc&Tur2012} developed two stopping rules that they use to estimate $K$. 
\citet{Duc&Tur2012} advocate running  {\sc flock} until either a single plateau of
$\geq 6$ runs is reached or four consecutive $K$ values yield no plateaus. If the first stopping condition is satisfied,
the $K$ for which there is a single plateau of $\geq 6$ runs is the point estimate for $K$. Alternatively, 
if multiple plateaus are observed with at least one plateau of $\geq 6$ runs, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. If no plateaus of $\geq 6$ runs are observed, $K$
is declared {\em undecided}.

@@

Once we were confident that no major differences in inference existed between  {\sc flock} and {\sc flockture}
we used  {\sc flockture} to analyze 48 datasets for two marker types 
(microsatellites and single nucleotide polymorphisms (SNPs))  that encompassed a broader 
range of population differentiation. The 96 datasets were simulated as above (five populations 
comprised of 250, 250, 400, 550, and 550 individuals with 15 loci genotypes for the micro satellites
and 96 loci for the SNPs)
with average pairwise $F_{ST}$ values ranging between 0.005 and 0.040. 
 {\sc flockture} was run with the default run parameters and each run of 
{\sc structure} consisted of a 5,000 sweeep burn-in followed by a 20,000 sweep sample
from the posterior distribution. All runs were performed on a Macbook with a 2.3 GHz i7 processor
and 16GB of RAM running OS X ver. 10.7.5. The program {\sc clumpp} \citep{Jak&Ros2007} 
was used to relabel clusters among different runs,
while the program {\sc district} \citep{Rosenberg2004} was used to plot individual \textit{$q_i$} values.

@@

  \begin{figure*}
\centering
\includegraphics[width=.9\linewidth]{images/Figures-Pat/uSatloss.pdf}%Need to make a greyscale version
  \caption{Zero-one loss over 9 replicate runs of the admixture and correlated allele 
  frequency (A.C) model, 
the no-admixture and correlated allele frequency (NA.C) model, 
and the no-admixture and non-correlated allele frequency (NA.NC) model in {\sc structure} and {\sc flockture}.}
  \label{fig:uSatLoss}
\end{figure*}

 \begin{figure*}
\centering
  \includegraphics[width=.9\linewidth]{images/Figures-Pat/SNPloss.pdf}%Need to make a greyscale version
  \caption{Zero-one loss over 9 replicate runs of the admixture and correlated allele 
  frequency (A.C) model, 
the no-admixture and correlated allele frequency (NA.C) model, 
and the no-admixture and non-correlated allele frequency (NA.NC) model in {\sc structure} and {\sc flockture}.}
  \label{fig:SNPloss}
\end{figure*}
